Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2014-10-29 MS_orig.tex
work on diff

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{hyperref}

% line numbers
\usepackage{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
%\usepackage{rotating}

% If you wish to include algorithms, please use one of the packages below. Also, please see the algorithm section of our LaTeX guidelines (http://www.plosone.org/static/latexGuidelines) for important information about required formatting.
%\usepackage{algorithmic}
%\usepackage{algorithmicx}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

\author{}
\date{}
\pagestyle{myheadings}

\begin{document}
% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Diversity dynamics in Nymphalidae butterflies: Effect of
phylogenetic uncertainty on diversification rate shift estimates}
}
% Insert Author names, affiliations and corresponding author email.
\\
Carlos Pe\~na$^{1,\ast}$, 
Marianne Espeland$^{2}$
\\
\bf{1} Carlos Pe\~na Laboratory of Genetics, Department of Biology, University of Turku, Turku
\\
\bf{2} Museum of Comparative Zoology and Department of Organismic and
Evolutionary Biology, Harvard University, Cambridge, USA. Email:
marianne.espeland@gmail.com.
Current address: McGuire Center for Lepidoptera and Biodiversity,
Florida Museum of Natural History, University of Florida, Gainesville,
FL, US
\\
$\ast$ E-mail: Corresponding mycalesis@gmail.com
\end{flushleft}

\section*{Abstract}

@@

\section*{Materials and Methods}

\subsection*{Data}

@@

\subsection*{Analyses of Diversification}

@@

\subsection*{Detecting diversification shifts on phylogenetic trees}

Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
version 093.4.33 {[}13{]} on the maximum clade credibility tree from
Reference 10. MEDUSA fits alternative birth-death likelihood models to a
phylogenetic tree in order to estimate changes in net diversification
rates along branches. MEDUSA estimates likelihood and AIC scores for the
simplest birth-death model, with two parameters, the rates \texttt{r}:
net diversification \texttt{=} speciation (\texttt{b}) \texttt{-}
extinction (\texttt{d}) and $\varepsilon$: relative extinction
\texttt{= d/b}. The AIC scores of the two-parameter model are then
compared with incrementally more complex models until the addition of
parameters do not improve the AIC scores beyond a cutoff value. MEDUSA
finds the likelihood of the models after taking into account branch
lengths and number of species per lineage {[}13{]}. Most studies using
MEDUSA only run the method on a single tree, usually the maximum clade
credibility tree, which makes the assumption that this tree is correct.
We wanted to study the effects of phylogenetic uncertainty on estimation
of net diversification rate shifts and therefore selected 1000 random
genus-level trees from the posterior distribution (MultiMEDUSA, supp.
mat. 5). We calculated a new MCC tree from the selection of trees, ran
MEDUSA and MultiMEDUSA on the selection of 1000 trees and summarized the
estimated changes in net diversification rates for nodes across all
trees. Patterns of change in net diversification rates are considered
significant if they are found at the same node in at least 90\% of the
trees. We also expected to find similar values of net diversification
(\texttt{r = b - d}) and relative extinction rate ($\varepsilon$
\texttt{= d/b}) values across the 1000 trees for the nodes where changes
in diversification tempo occurs. We used the AICc threshold of 7.8
units, as estimated by MEDUSA, as the limit for a significantly better
fit to select among increasingly complex alternative models.

\subsection*{Estimation of trait-dependent speciation rates}

@@

\section*{Results}

\subsection*{Detecting diversification shifts on the maximum clade credibility tree}

@@

\subsection*{Phylogenetic uncertainty in the MultiMEDUSA approach}

@@

\subsection*{Estimation of trait-dependent speciation rates}

@@

\section*{Discussion}

\subsection*{Effects of phylogenetic uncertainty on the performance of MEDUSA}

@@

\subsection*{Hostplant use and diversification in Nymphalidae}

\subsubsection*{Ithomiini}

@@

\subsubsection*{Danaini}

@@

\subsubsection*{\emph{Charaxes}}

@@

\subsubsection*{Satyrini}

@@

\section*{Acknowledgments}

We are thankful to Mark Cornwall for help with the script to extend
MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
commenting on the manuscript and giving us the posterior distribution of
trees, Luke Harmon for commenting on the manuscript and anonymous
reviewers for their comments, which greatly improved the manuscript,
Jessica Slove Davidson and Niklas Janz for access to their hostplant
data. We acknowledge CSC--IT Center for
Science Ltd. (Finland) for the allocation of computational resources.

\section*{Figure legends}

\texttt{Figure 1.} MEDUSA analysis run on the MCT 
from Reference 10. Estimated rate shifts in Table 1.

\texttt{Figure 2.} Range of diversification values for
tips or clades estimated by MEDUSA on the 1000 random trees.

\texttt{Figure 3.} Results of the MultiMEDUSA analysis showing
the probability for nodes of being recovered as diversification events.

\texttt{Figure 4.} BiSSE analysis of diversification of Solanaceae feeding 
nymphalids shows higher speciation and net diversification
rates in Solanaceae feeders.

\section*{References}
1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
adaptive zones: has phytophagy promoted insect diversification? Am Nat
132: 107--128.
doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.

2. Nyman T, Linder HP, Pe\~na C, Malm T, Wahlberg N (2012) Climate-driven
diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
889--898.
doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.

@@

6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
al. (2011) Order Lepidoptera. In: Animal biodiversity: an outline of
higher-level classification and survey of taxonomic richness. Zootaxa.
pp. 212--221.

7. Heikkil\"a M, Kaila L, Mutanen M, Pe\~na C, Wahlberg N (2012) Cretaceous
origin and repeated tertiary diversification of the redefined
butterflies. P Roy Soc B 279: 1093--1099.
doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.

8. Elias M, Joron M, Willmott KR, Silva-Brand\~ao KL, Kaiser V, et al.
(2009) Out of the Andes: Patterns of diversification in clearwing
butterflies. Mol Ecol 18: 1716--1729.
doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.

@@

10. Wahlberg N, Leneveu J, Kodandaramaiah U, Pe\~na C, Nylin S, et al.
(2009) Nymphalid butterflies diversify following near demise at the
Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.

11. Pe\~na C, Wahlberg N (2008) Prehistorical climate change increased
diversification of a group of butterflies. Biol Lett 4: 274--278.
doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.

12. Ferrer-Paris JR, S\'anchez-Mercado A, Viloria \'AL, Donaldson J (2013)
Congruence and Diversity of Butterfly-Host Plant Associations at Higher
Taxonomic Levels. PloS one 8: e63570.
doi:\href{http://dx.doi.org/10.1371/journal.pone.0063570}{10.1371/journal.pone.0063570}.

@@

14. Harmon LJ, Rabosky DL, FitzJohn RG, Brown JW (2011) turboMEDUSA:
MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.

15. FitzJohn RG (2012) Diversitree: comparative phylogenetic analyses of
diversification in R. Method Ecol Evol 3: 1084--1092.
doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.

@@

17. Matos-Marav\'i PF, Pe\~na C, Willmott KR, Freitas AV, Wahlberg N (2013)
Systematics and evolutionary history of butterflies in the ``Taygetis
clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
54--68.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.

18. Brower AVZ, Wahlberg N, Ogawa JR, Boppr\'e M, Vane-Wright RI (2010)
Phylogenetic relationships among genera of danaine butterflies
(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
Syst Biodivers 8: 75--89.
doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.

19. Kodandaramaiah U, Lees DC, M\"uller CJ, Torres E, Karanth KP, et al.
(2010) Phylogenetics and biogeography of a spectacular Old World
radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
Nymphalidae: Satyrini). BMC Evol Biol 10: 172. 
doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.

20. Kodandaramaiah U, Pe\~na C, Braby MF, Grund R, M\"uller CJ, et al.
(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.

@@

24. Pe\~na C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.

@@

26. Silva-Brand\~ao KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.

27. Pe\~na C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
methods. Zool J Linn Soc 161: 64--87.
doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.

28. Pe\~na C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
Satyrinae). Zool Scr 39: 243--258.
doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.

29. Ackery PR (1988) Hostplants and classification: a review of
nymphalid butterflies. Biol J Linn Soc 33: 95--203.
doi:\href{http://dx.doi.org/10.1111/j.1095-8312.1988.tb00446.x}{10.1111/j.1095-8312.1988.tb00446.x}.

@@

31. Beccaloni G, Viloria \'A, Hall S, Robinson G (2008) Catalogue of the
hostplants of the Neotropical butterflies. London: The Natural History
Museum.

@@

43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brand\~ao KL, Whinnett A, et
al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
Nymphalidae) inferred from one mitochondrial and two nuclear gene
regions. Syst Entomol 31: 288--301.
doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.

44. FitzJohn RG, Maddison WP, Otto SP (2009) Estimating trait-dependent
speciation and extinction rates from incompletely resolved phylogenies.
Syst Biol 58: 595--611. Available:
doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.

@@

51. Wahlberg N, Rubinoff D (2011) Vagility across \textit{Vanessa} (Lepidoptera:
Nymphalidae): mobility in butterfly species does not inhibit the
formation and persistence of isolated sister taxa. Syst Entomol 36:
362--370.
doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.

52. M\"uller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
evolutionary history and systematics of the genus \textit{Charaxes} Ochsenheimer
(Lepidoptera: Nymphalidae) in the Indo-Pacific region. Biol J Linn Soc
100: 457--481.
doi:\href{http://dx.doi.org/10.1111/j.1095-8312.2010.01426.x}{10.1111/j.1095-8312.2010.01426.x}.

>>>>

\documentclass[]{article}
\usepackage{setspace}
\usepackage[]{lineno}
\def\linenumberfont{\normalfont\small\sffamily}
\doublespacing
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage[none]{hyphenat}
\usepackage{fixltx2e} % provides \textsubscript
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
\usepackage[margin=1in,a4paper]{geometry}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\linenumbers

\author{}
\date{}

\begin{document}

\textbf{Title: Diversity dynamics in Nymphalidae butterflies: Effect of
phylogenetic uncertainty on diversification rate shift estimates}

\subsubsection{Authors: Carlos Peña\textsuperscript{1,*} and Marianne
Espeland\textsuperscript{2,*}}\label{authors-carlos-peuxf1a1-and-marianne-espeland2}

\subsubsection{Running title: Diversity dynamics in Nymphalidae
butterflies}\label{running-title-diversity-dynamics-in-nymphalidae-butterflies}

\emph{1 Laboratory of Genetics, Department of Biology, University of
Turku, Turku, Finland. Email:
\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}}

\emph{2 Museum of Comparative Zoology and Department of Organismic and
Evolutionary Biology, Harvard University, Cambridge, USA. Email:
\href{mailto:marianne.espeland@gmail.com}{\nolinkurl{marianne.espeland@gmail.com}}.
Current address: McGuire Center for Lepidoptera and Biodiversity,
Florida Museum of Natural History, University of Florida, Gainesville,
FL, US}.

\textbf{Corresponding author:}
\href{mailto:mycalesis@gmail.com}{\nolinkurl{mycalesis@gmail.com}}
Telephone: +358 417063065 Fax: +358 2 333 6680

\textbf{Comments:} 5814 words, 7 figures, 2 tables and 25 supplementary
material files.

\textbf{Additional Supplementary materials:}
\url{http://dx.doi.org/10.6084/m9.figshare.639208}

\subsection{Abstract}\label{abstract}

@@

\subsection{Methods}\label{methods}

\subsubsection{Data}\label{data}

@@

\subsubsection{Analyses of
Diversification}\label{analyses-of-diversification}

@@

\subsubsection{Detecting diversification shifts on phylogenetic
trees}\label{detecting-diversification-shifts-on-phylogenetic-trees}

Patterns of diversification in Nymphalidae were analyzed by using MEDUSA
version 093.4.33 {[}13{]} on the maximum clade credibility tree from
Reference 10. MEDUSA fits alternative birth-death likelihood models to a
phylogenetic tree in order to estimate changes in net diversification
rates along branches. MEDUSA estimates likelihood and AIC scores for the
simplest birth-death model, with two parameters, the rates \texttt{r}:
net diversification \texttt{=} speciation (\texttt{b}) \texttt{-}
extinction (\texttt{d}) and \(\varepsilon\): relative extinction
\texttt{= d/b}. The AIC scores of the two-parameter model are then
compared with incrementally more complex models until the addition of
parameters do not improve the AIC scores beyond a cutoff value. MEDUSA
finds the likelihood of the models after taking into account branch
lengths and number of species per lineage {[}13{]}. Most studies using
MEDUSA only run the method on a single tree, usually the maximum clade
credibility tree, which makes the assumption that this tree is correct.
We wanted to study the effects of phylogenetic uncertainty on estimation
of net diversification rate shifts and therefore selected 1000 random
genus-level trees from the posterior distribution (MultiMEDUSA, supp.
mat. 5). We calculated a new MCC tree from the selection of trees, ran
MEDUSA and MultiMEDUSA on the selection of 1000 trees and summarized the
estimated changes in net diversification rates for nodes across all
trees. Patterns of change in net diversification rates are considered
significant if they are found at the same node in at least 90\% of the
trees. We also expected to find similar values of net diversification
(\texttt{r = b - d}) and relative extinction rate (\(\varepsilon\)
\texttt{= d/b}) values across the 1000 trees for the nodes where changes
in diversification tempo occurs. We used the AICc threshold of 7.8
units, as estimated by MEDUSA, as the limit for a significantly better
fit to select among increasingly complex alternative models.

\subsubsection{Estimation of trait-dependent speciation
rates}\label{estimation-of-trait-dependent-speciation-rates}

@@

\subsection{Results}\label{results}

\subsubsection{Detecting diversification shifts on the maximum clade
credibility
tree}\label{detecting-diversification-shifts-on-the-maximum-clade-credibility-tree}

@@

\subsubsection{Phylogenetic uncertainty in the MultiMEDUSA
approach}\label{phylogenetic-uncertainty-in-the-multimedusa-approach}

@@

\subsubsection{Estimation of trait-dependent speciation
rates}\label{estimation-of-trait-dependent-speciation-rates-1}

@@

\subsection{Discussion}\label{discussion}

\subsubsection{Effects of phylogenetic uncertainty on the performance of
MEDUSA}\label{effects-of-phylogenetic-uncertainty-on-the-performance-of-medusa}

@@

\subsubsection{Hostplant use and diversification in
Nymphalidae}\label{hostplant-use-and-diversification-in-nymphalidae}

\paragraph{Ithomiini}\label{ithomiini}

@@

\paragraph{Danaini}\label{danaini}

@@

\paragraph{\emph{Charaxes}}\label{charaxes}

@@

\paragraph{Satyrini}\label{satyrini}

@@

\subsection{Acknowledgments}\label{acknowledgments}

We are thankful to Mark Cornwall for help with the script to extend
MEDUSA to include phylogenetic uncertainty, Niklas Wahlberg for
commenting on the manuscript and giving us the posterior distribution of
trees, Luke Harmon for commenting on the manuscript and anonymous
reviewers for their comments, which greatly improved the manuscript,
Jessica Slove Davidson and Niklas Janz for access to their hostplant
data. The study was supported by a Kone Foundation grant (awarded to
Niklas Wahlberg), Finland (C. Peña) and the Research Council of Norway
(grant no. 204308 to M. Espeland). We acknowledge CSC--IT Center for
Science Ltd. (Finland) for the allocation of computational resources.

\subsection{Figure legends}\label{figure-legends}

\texttt{Figure 1.} Results of the MEDUSA analysis run on the maximum
clade credibility tree from Reference 10. Rate shifts were estimated for
the following nodes (besides the background rate): 2) Limenitidinae +
Heliconiinae, 3) \emph{Ypthima}, 4) \emph{Charaxes}, 5) Ithomiini in
part, 6) Satyrini, 7) Coenonymphina 8) Phyciodina in part, 9) Danaini in
part, 10) \emph{Coenonympha}, 11) \emph{Caeruleuptychia} +
\emph{Magneuptychia}, 12) \emph{Callicore} + \emph{Diaethria}, 13)
Satyrina, 14) Mycalesinaa, 15) \emph{Pedaliodes}, 16) \emph{Dryas} +
\emph{Dryadula}, 17) \emph{Taenaris}, 18) Pseudergolinae.

\texttt{Figure 2.} Boxplot of the range of diversification values for
tips or clades estimated by MEDUSA on the 1000 random trees from the
posterior distribution of the Nymphalidae phylogeny. The tips or clades
shown are those present on the maximum clade credibility tree from
Reference 10.

\texttt{Figure 3.} Results of the MultiMEDUSA analysis on 1000 randomly
sampled trees from the posterior distribution of the Nymphalidae
phylogeny. Bars show the probability for nodes of being recovered as
significant increases or decreases in net diversification rates by
MultiMEDUSA.

\texttt{Figure 4.} BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants. Speciation and net diversification
rates are significantly higher in Solanaceae feeders (\$lambda\$1, r1).

\subsection{Tables}\label{tables}

\texttt{Table 1.} Significant net diversification rate shifts found in
the MEDUSA analysis of the Nymphalid maximum clade credibility tree.
Split.Node = node number, Model = preferred diversification model by
MEDUSA, r = net diversification rate, LnLik.part = log likelihood value.

\texttt{Table 2.} Differences in rates estimated by MEDUSA on the MCC
tree from Wahlberg and the MultiMEDUSA approach on 1000 random trees
from the posterior distribution.

\texttt{Table 3.} Likelihood ratio test between the model of increased
diversification of nymphalids feeding on Solanaceae against a model
forcing equal speciation rates (no effect on diversification).

\section{Supp. mat.}\label{supp.-mat.}

\textbf{\texttt{supp\_mat\_01\_genus.nex}}: MCC Nymphalidae tree from
Reference 10.

\textbf{\texttt{supp\_mat\_02\_run\_medusa\_on\_mct.R}}: Run MEDUSA on
MCC tree from Reference 10.

\textbf{\texttt{supp\_mat\_03\_richness.nex}}: Species richness for
lineages in Nymphalidae.

\textbf{\texttt{supp\_mat\_04\_MEDUSA\_results.txt}}: Results of running
MEDUSA on the MCC tree.

\textbf{\texttt{supp\_mat\_05\_1000\_random\_trees\_no\_outgroups.nex}}:
1000 random trees from Reference 10.

\textbf{\texttt{supp\_mat\_06\_run\_multimedusa.R}}: Run MultiMEDUSA on
1000 random trees from Reference 10.

\textbf{\texttt{supp\_mat\_07\_multimedusa\_on\_1000\_trees.txt}}: Raw
results from the MultiMEDUSA run on the random sample of trees from the
posterior distribution.

\textbf{\texttt{supp\_mat\_08\_multimedusa\_result01.pdf}}: Summary of
MulitMEDUSA analysis on the 1000 trees from the posterior distribution.

\textbf{\texttt{supp\_mat\_09\_mcc\_tree\_from\_random\_sample\_of\_1000.tree}}:
MCC tree from the 1000 random trees selected from the posterior
distribution.

\textbf{\texttt{supp\_mat\_10\_run\_medusa\_on\_mct\_from\_1000\_trees.R}}:
Run MEDUSA on MCC tree from the 1000 random trees selected from the
posterior distribution.

\textbf{\texttt{supp\_mat\_11\_results\_of\_MEDUSA\_run\_on\_mcc\_from\_1000\_trees.txt}}:
Summary results from a MEDUSA run on the MCC tree from the 1000 random
trees selected from the posterior distribution.

\textbf{\texttt{supp\_mat\_12\_MEDUSA\_mcc\_tree\_from\_random\_sample\_of\_1000\_trees.pdf}}:
Figure for MEDUSA run on MCC tree from random 1000 trees.

\textbf{\texttt{supp\_mat\_13\_results\_of\_Multi-MEDUSA\_run\_on\_1000\_trees.txt}}:
Summary results from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution.

\textbf{\texttt{supp\_mat\_14\_multimedusa\_result\_prob\_nodes.pdf}}:
Probability of nodes with estimated rates from a MultiMEDUSA run on the
1000 random trees selected from the posterior distribution.

\textbf{\texttt{supp\_mat\_15\_host\_plant\_data.csv}}: Hostplants of
Nymphalidae butterflies recorded from the literature.

\textbf{\texttt{supp\_mat\_16\_host\_plant\_data\_references.csv}}:
References for hostplants data.

\textbf{\texttt{supp\_mat\_17\_hostplant\_states.csv}}: Data matrix with
character states for hosplant use.

\textbf{\texttt{supp\_mat\_18\_bisse\_source.R}}: R code for running the
BiSSE analysis.

\textbf{\texttt{supp\_mat\_19\_bisse\_mcmc\_run.csv}}: Raw results for
the BiSSE analysis.

\textbf{\texttt{supp\_mat\_20\_combined\_bisse.csv}}: Raw results for
the combined BiSSE analysis.

\texttt{Figure S1.} Boxplot of speciation (lambda) and extinction (mu)
values for the 95\% credibility intervals of values estimated by BiSSE
analysis of diversification due to feeding on Solanaceae plants.

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

1. Mitter C, Farrell B, Wiegmann B (1988) The phylogenetic study of
adaptive zones: Has phytophagy promoted insect diversification? Am Nat
132: 107--128.
doi:\href{http://dx.doi.org/10.1086/284840}{10.1086/284840}.

2. Nyman T, Linder HP, Peña C, Malm T, Wahlberg N (2012) Climate-driven
diversity dynamics in plants and plant-feeding insects. Ecol Lett 15:
889--898.
doi:\href{http://dx.doi.org/10.1111/j.1461-0248.2012.01782.x}{10.1111/j.1461-0248.2012.01782.x}.

@@

6. Nieukerken EJ van, Kaila L, Kitching IJ, Kristensen NP, Lees DC, et
al. (2011) Order Lepidoptera. In: Animal biodiversity: An outline of
higher-level classification and survey of taxonomic richness. Zootaxa.
pp. 212--221.

7. Heikkilä M, Kaila L, Mutanen M, Peña C, Wahlberg N (2012) Cretaceous
origin and repeated tertiary diversification of the redefined
butterflies. P Roy Soc B 279: 1093--1099.
doi:\href{http://dx.doi.org/10.1098/rspb.2011.1430}{10.1098/rspb.2011.1430}.

8. Elias M, Joron M, Willmott KR, Silva-Brandão KL, Kaiser V, et al.
(2009) Out of the Andes: Patterns of diversification in clearwing
butterflies. Mol Ecol 18: 1716--1729.
doi:\href{http://dx.doi.org/10.1111/j.1365-294X.2009.04149.x}{10.1111/j.1365-294X.2009.04149.x}.

@@

10. Wahlberg N, Leneveu J, Kodandaramaiah U, Peña C, Nylin S, et al.
(2009) Nymphalid butterflies diversify following near demise at the
Cretaceous/Tertiary boundary. P Roy Soc B 276: 4295--4302.
doi:\href{http://dx.doi.org/10.1098/rspb.2009.1303}{10.1098/rspb.2009.1303}.

11. Peña C, Wahlberg N (2008) Prehistorical climate change increased
diversification of a group of butterflies. Biol Lett 4: 274--278.
doi:\href{http://dx.doi.org/10.1098/rsbl.2008.0062}{10.1098/rsbl.2008.0062}.

12. Ferrer-Paris JR, S{á}nchez-Mercado A, Viloria ÁL, Donaldson J (2013)
Congruence and Diversity of Butterfly-Host Plant Associations at Higher
Taxonomic Levels. PloS one 8: e63570.
doi:\href{http://dx.doi.org/0.1371/journal.pone.0063570}{0.1371/journal.pone.0063570}.

@@

14. Harmon LJ, Rabosky DL, {FitzJohn} RG, Brown JW (2011) turboMEDUSA:
MEDUSA: Modeling Evolutionary Diversification Using Stepwise AIC.

15. {FitzJohn} RG (2012) Diversitree: comparative phylogenetic analyses
of diversification in R. Method Ecol Evol 3: 1084--1092.
doi:\href{http://dx.doi.org/10.1111/j.2041-210X.2012.00234.x}{10.1111/j.2041-210X.2012.00234.x}.

@@

17. Matos-Maraví PF, Peña C, Willmott KR, Freitas AV, Wahlberg N (2013)
Systematics and evolutionary history of butterflies in the ``Taygetis
clade'' (Nymphalidae: Satyrinae: Euptychiina): Towards a better
understanding of Neotropical biogeography. Mol Phylogenet Evol 66:
54--68.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2012.09.005}{10.1016/j.ympev.2012.09.005}.

18. Brower AVZ, Wahlberg N, Ogawa JR, Boppré M, Vane-Wright RI (2010)
Phylogenetic relationships among genera of danaine butterflies
(Lepidoptera: Nymphalidae) as implied by morphology and DNA sequences.
Syst Biodivers 8: 75--89.
doi:\href{http://dx.doi.org/10.1080/14772001003626814}{10.1080/14772001003626814}.

19. Kodandaramaiah U, Lees DC, Müller CJ, Torres E, Karanth KP, et al.
(2010) Phylogenetics and biogeography of a spectacular Old World
radiation of butterflies: the subtribe Mycalesina (Lepidoptera:
Nymphalidae: Satyrini). BMC Evol Biol 10: 172.
doi:\href{http://dx.doi.org/10.1186/1471-2148-10-172}{10.1186/1471-2148-10-172}.

20. Kodandaramaiah U, Peña C, Braby MF, Grund R, Müller CJ, et al.
(2010) Phylogenetics of Coenonymphina (Nymphalidae: Satyrinae) and the
problem of rooting rapid radiations. Mol Phylogenet Evol 54: 386--394.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2009.08.012}{10.1016/j.ympev.2009.08.012}.

@@

24. Peña C, Wahlberg N, Weingartner E, Kodandaramaiah U, Nylin S, et al.
(2006) Higher level phylogeny of Satyrinae butterflies (Lepidoptera:
Nymphalidae) based on DNA sequence data. Mol Phylogenet Evol 40: 29--49.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2006.02.007}{10.1016/j.ympev.2006.02.007}.

@@

26. Silva-Brandão KL, Wahlberg N, Francini RB, Azeredo-Espin AML, Brown
Jr. KS, et al. (2008) Phylogenetic relationships of butterflies of the
tribe Acraeini (Lepidoptera, Nymphalidae, Heliconiinae) and the
evolution of host plant use. Mol Phylogenet Evol 46: 515--531.
doi:\href{http://dx.doi.org/10.1016/j.ympev.2007.11.024}{10.1016/j.ympev.2007.11.024}.

27. Peña C, Nylin S, Wahlberg N (2011) The radiation of Satyrini
butterflies (Nymphalidae: Satyrinae): a challenge for phylogenetic
methods. Zool J Linn Soc 161: 64--87.
doi:\href{http://dx.doi.org/10.1111/j.1096-3642.2009.00627.x}{10.1111/j.1096-3642.2009.00627.x}.

28. Peña C, Nylin S, Freitas AVL, Wahlberg N (2010) Biogeographic
history of the butterfly subtribe Euptychiina (Lepidoptera, Nymphalidae,
Satyrinae). Zool Scr 39: 243--258.
doi:\href{http://dx.doi.org/10.1111/j.1463-6409.2010.00421.x}{10.1111/j.1463-6409.2010.00421.x}.

29. Ackery PR (1988) Hostplants and classification: A review of
nymphalid butterflies. Biol J Linn Soc 33: 95--203.
doi:\href{http://dx.doi.org/10.1111/j.1095-8312.1988.tb00446.x}{10.1111/j.1095-8312.1988.tb00446.x}.

@@

31. Beccaloni G, Viloria Á, Hall S, Robinson G (2008) Catalogue of the
hostplants of the Neotropical butterflies. London: The Natural History
Museum.

@@

43. Brower AVZ, Freitas AVL, Lee M-M, Silva-Brandão KL, Whinnett A, et
al. (2006) Phylogenetic relationships among the Ithomiini (Lepidoptera:
Nymphalidae) inferred from one mitochondrial and two nuclear gene
regions. Syst Entomol 31: 288--301.
doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2006.00321.x}{10.1111/j.1365-3113.2006.00321.x}.

44. {FitzJohn} RG, Maddison WP, Otto SP (2009) Estimating
trait-dependent speciation and extinction rates from incompletely
resolved phylogenies. Syst Biol 58: 595--611.
doi:\href{http://dx.doi.org/10.1093/sysbio/syp067}{10.1093/sysbio/syp067}.

@@

51. Wahlberg N, Rubinoff D (2011) Vagility across Vanessa (Lepidoptera:
Nymphalidae): mobility in butterfly species does not inhibit the
formation and persistence of isolated sister taxa. Syst Entomol 36:
362--370.
doi:\href{http://dx.doi.org/10.1111/j.1365-3113.2010.00566.x}{10.1111/j.1365-3113.2010.00566.x}.

52. Müller CJ, Wahlberg N, Beheregaray LB (2010) `After Africa': the
evolutionary history and systematics of the genus Charaxes Ochsenheimer
(Lepidoptera: Nymphalidae) in the Indo-Pacific region. Biol J Linn Soc
100: 457--481.
doi:\href{http://dx.doi.org/10.1111/j.1095-8312.2010.01426.x}{10.1111/j.1095-8312.2010.01426.x}.
